Constrained Optimization-Based Neuro-Adaptive Control (CONAC) for Euler-Lagrange Systems Under Weight and Input Constraints

### 2025-06-08 manuscript.tex
final touch for shJang

\author{
  Myeongseok Ryu, Donghwa Hong, and Kyunghwan Choi
\thanks{
    Myeongseok Ryu and Kyunghwan Choi are with the Cho Chun Shik Graduate School of Mobility, Korean Advanced Institute of Science and Technology (KAIST), Daejeon 34051, Korea (e-mail: dding\_98@kaist.ac.kr and kh.choi@kaist.ac.kr).

@@

%% Abstract
\begin{abstract}
  This study presents a constrained optimization-based neuro-adaptive controller (CONAC) for uncertain Euler-Lagrange systems subject to weight norm and convex input constraints. 
  A deep neural network (DNN) is employed to approximate the ideal stabilizing control law while simultaneously learning the unknown system dynamics and addressing both types of constraints within a constrained optimization framework.
  The weight adaptation laws are formulated to solve the defined constrained optimization problem, ensuring satisfaction of first-order optimality conditions at steady state.
  The controller's stability is rigorously analyzed using Lyapunov theory, guaranteeing bounded tracking errors and bounded DNN weights. 
  The proposed controller is validated through a real-time implementation on a 2-DOF robotic manipulator, demonstrating its effectiveness in achieving accurate trajectory tracking while satisfying all imposed constraints.
\end{abstract}

@@

\begin{itemize}
    \item $\otimes$ denotes the Kronecker product \cite[Chap. 7 Def. 7.1.2]{Bernstein:2009aa}.
    \item A vector and a matrix are denoted by $\mv{x}=[x_i]_{i\in\{1,\cdots,n\}}\in\R^n$ and $
        \mm A
        := 
        [a_{ij}]
        _{
            i\in\{1,\cdots,n\},j\in\{1,\cdots ,m\}
        }\allowbreak\in\R^{n\times m}
        $, respectively.
    \item $\myrow_i(\mm A)$ denotes the $i\textsuperscript{th}$ row of the matrix $\mm A\in\R^{n\times m}$. 
    \item For $\mm{A}\in\R^{n\times m}$, $\myvec(\mm A):=(\myrow_1(\mm A^\top),\cdots,\myrow_m(\mm A^\top)  )^\top\in\R^{nm}$ denotes the vectorization of $\mm A$.
    % \item $\mycol_i(A)$ and $\myrow_j(A)$ denote the $i\textsuperscript{th}$ column and $j\textsuperscript{th}$ row of $A\in\R^{n\times m}$, respectively.
    % \item $\myvec(A):= [\mycol_1(A)^\top  ,\cdots,\mycol_m(A)^\top  ]^\top   $ for $A\in\R^{n\times m}$.
    \item $\lambda_{\min}(\mm A)$ denotes the minimum eigenvalue of the matrix $\mm A\in\R^{n\times n}$.
    \item $\mm I_n$ denotes the $n\times n$ identity matrix, and $\mm 0_{n\times m}$ denotes the $n\times m$ zero matrix.
\end{itemize}

@@

More recently, neuro-adaptive control approaches have been introduced to approximate unknown system dynamics or entire control laws using neural networks (NNs) \cite{Farrell:2006aa}. 
NNs are well-known for their universal approximation property, which allows them to approximate any smooth function over a compact set with minimal error. 
Various types of NNs have been utilized in neuro-adaptive control, including simpler architectures like single-hidden layer (SHL) NNs \cite{Ge:2010aa, Yesildirek:1995aa} and radial basis function (RBF) NNs \cite{Liu:2013ab,Ge:2002aa}, as well as more complex models like deep NNs (DNNs) \cite{Patil:2022aa} and their variations. 
SHL and RBF NNs are often employed to approximate uncertain system dynamics or controllers due to their simplicity \cite{Esfandiari:2014aa,Esfandiari:2015aa,Yesildirek:1995aa,Gao:2006aa}.
However, since DNNs offer greater expressive power, making them more effective for complex system approximations \cite{Rolnick:2018aa}, variations of DNNs, such as long short-term memory (LSTM) networks for time-varying dynamics \cite{Griffis:2023aa} and physics-informed NNs (PINNs) for leveraging physical system knowledge \cite{Hart:2024aa}, have further extended the capabilities of neuro-adaptive control systems.

A critical aspect of neuro-adaptive control is the weight adaptation law, which governs how NN weights are updated. 
Most studies derived these laws using Lyapunov-based methods, ensuring the boundedness of the tracking error and weight estimation error, thus maintaining system stability under uncertainty.

@@

Although both the projection operator methods for weight norm boundedness and the auxiliary system approach for input constraints have shown effectiveness, they come with significant limitations. 
Projection operators and modification techniques do not guarantee the optimality of the adapted weights. 
Moreover, auxiliary systems typically handle only simple forms of input constraints, such as input bound constraints, limiting their ability to address more complex, nonlinear constraints.

@@

The main contributions of this study are listed as follows:
\begin{itemize}
    \item A constrained optimization-based neuro-adaptive controller (CONAC) is proposed, where the design of a deep neural network (DNN) is formulated as a constrained optimization problem. In this formulation, weight and convex input constraints are incorporated as inequality constraints, while minimizing a given objective function.
    % \item Any convex input constraints can be applied to the proposed CONAC, including input bound constraints, and norm constraints. Moreover, the proposed CONAC can handle any combination of these constraints, which can be useful in practical applications where multiple constraints need to be satisfied simultaneously.
    % \item Since the proposed CONAC approximate the entire ideal control law with no prior knowledge of the system dynamics, the system identification process can be avoided, which is often difficult and time-consuming to implement in practice. 
    \item The adaptation laws for DNN weights in proposed CONAC are systematically derived to solve the defined problem, using constrained optimization theory. These laws guarantee convergence to the first-order optimality conditions, specifically the Karush–Kuhn–Tucker (KKT) conditions, as described in \cite[Chap. 12 Thm. 12.1]{Nocedal:2006aa}.
    \item The proposed CONAC's stability is rigorously analyzed using Lyapunov theory, ensuring that the tracking error and DNN weights remain bounded. This analysis guarantees that the proposed CONAC can achieve stable adaptation online.
    \item The proposed CONAC is implemented in real-time on a 2-DOF robotic manipulator, demonstrating its effectiveness in achieving accurate trajectory tracking while satisfying all imposed constraints. The experimental results validate the theoretical findings and show the practical applicability of the proposed controller.
\end{itemize}

@@

The remainder of this paper is organized as follows. 
Section \ref{sec:Problem Formulation} presents the target system and control objective.
Section \ref{sec:ctrl design} introduces the proposed controller and the architecture of DNN in the controller. 
In Section \ref{sec:adap_laws}, the weight adaptation laws are developed.
Section \ref{sec:stability} analyzes the stability of the proposed controller.
A real-time implementation of the proposed controller is presented in Section \ref{sec:sim}, where the proposed controller is applied to a 2-DOF robotic manipulator.
Finally, Section \ref{sec:conclusion} concludes the paper and discusses potential future work.
The candidates of input constraints are presented in Appendix \ref{sec:appen:cstr}.

@@

We consider an uncertain Euler-Lagrange system modeled as
\begin{equation}
    \rbM\ddq + \rbVm + \rbF + \rbG + \rbd
    =
    \mysat(\rbu)
    \label{eq:sys1}
\end{equation}
where $\q\in \R^n$ denotes the generalized coordinate, $\rbd$ represents the disturbance and $\rbu\in\R^n$ denotes the generalized control input. 
The terms $\rbM:=\rbM(\q)\in\R^{n\times n}$, $\rbVm:=\rbVm(\q,\dq)\in\R^{n\times n}$, $\rbF:=\rbF(\dq)\in\R^{n}$, and $\rbG:=\rbG(\q)$ represent the unknown inertia matrix, Coriolis/centripetal matrix, friction vector, and gravity vector, respectively.
The function $\mysat(\cdot):\R^n\to\R^n$ represents the inherent physical limitations of the actuators such that $\norm{\mysat(\rbu)}\le \overline\tau$, where $\overline\tau$ denotes the maximum norm of control input.
In this study, the saturation function $\mysat(\cdot)$ is assumed to be convex, which is practically common in many engineering systems, since the actuator limits are naturally convex, such as min/max torque limits, and bounds look like a box or ball in the input space.
To account for these limitations, it is essential to incorporate physically motivated constraints into the controller design.
Appendix \ref{sec:appen:cstr} introduces candidate constraints that can be applied to ensure compliance with these physical limitations.

@@

\begin{prop}
    The disturbance $\rbd$ is bounded, \ie $\norm{\rbd}\le \overline\tau_d$.
    \label{prop:dis_bound}
\end{prop}

The control objective is to develop a neuro-adaptive controller that enables $\q$ to track a continuously differentiable desired trajectory $\qd:=\qd(t): \R_{\ge0} \to \R^n$, compensating for the unknown system dynamics while addressing the imposed constraints.
The desired trajectory $\qd(t)$ is supposed to satisfy the following assumption:
\begin{assum}
    The desired trajectory $\qd(t)$ is assumed to be bounded, i.e., $\norm{\qd(t)}\le \overline{q}_d$ , and available for designing a bounded control input.
    \label{assum:feasible}
\end{assum}

@@

The architecture of the proposed CONAC is illustrated in Fig.~\ref{fig:ctrl:diagram}, and consists of a DNN that functions as a neuro-adaptive controller, and a weight optimizer for the DNN.
Before proceeding to the controller design, some mathematical preliminaries are revisited in Section \ref{sec:sub:math preliminaries}. 
Section \ref{sec:sub:NAC} introduces the neuro-adaptive controller, followed by the DNN model in Section \ref{sec:sub:NN definition}. 
The constrained optimization-based weight adaptation laws are presented in Section \ref{sec:adap_laws}.

@@

\begin{proof}
    The time derivative of $V$ can be rewritten as
    \begin{equation}
        \ddtt V
        \le
        -a_1 \norm{\mv{x}}^2 + a_2 \norm{\mv{x}}
        =
        -a_1
        \left(
            \norm{\mv{x}}-\tfrac{a_2}{2a_1}
        \right)^2
        -
        \tfrac{a_2^2}{4a_1}
        .
        \label{eq:lem:lya:dot}
    \end{equation}
    According to \eqref{eq:lem:lya:dot}, one can conclude that $\ddtt V$ is negative definite if $\norm{\mv{x}}\le\tfrac{a_2}{a_1}$ holds.
    In other words, $\mv{x}$ remains within the bounded set $\{\mv{x}\mid\norm{\mv{x}}\le\tfrac{a_2}{a_1}\}$.
\end{proof}

@@

First, the filtered tracking error $\fe\in\R^n$ is defined as 
\begin{equation}
    \fe := \ddtt{\mv e}-\mm\Lambda \mv e
    ,
    \label{eq:err:filtered}
\end{equation}
where $\mv e := \q - \qd$ denotes the tracking error, $\ddtt{\mv e} := \dq - \dqd$ represents the time derivative of the tracking error, and $\mm\Lambda\in\R^{n\times n}_{>0}$ is a user-designed filtering matrix.
Since \eqref{eq:err:filtered} is stable system, it implies that $\mv e$ is bounded if $\fe$ is bounded.

Using $\fe$, the system dynamics \eqref{eq:sys1} can be rewritten as
\begin{equation}
    \rbM \ddtt\fe
    =
    -\rbVm \fe
    -\mm K \fe
    + \mv f
    -\rbd + \mysat(\rbu)
    ,
    \label{eq:sys:filtered}
\end{equation}
where $\mm{K}=\mm{K}^\top\in\R^{n\times n}_{>0}$ denotes an arbitrary unknown matrix and $
    \mv f:= \mv f(\q,\dq,\qd,\dqd,\ddqd)
    =
    \mm K \fe
    +\rbM(-\ddqd+\mm\Lambda\ddtt{\mv e})
    +
    \rbVm(-\dqd+\mm\Lambda\mv e)
    -
    \rbF
    -
    \rbG
    \in\R^n
$ denotes the lumped system uncertainty.
% \color{red}
% The unknown matrix $\mm{K}$ is not available to design in practice, 
% \color{black}

Consider the Lyapunov function $V_1:= \tfrac{1}{2}\fe^\top \rbM\fe$. 
Invoking Property \ref{prop:skew}, the time derivative of $V_1$ is
\begin{equation}
    \begin{aligned}
        \ddtt{V}_1
        =&
        \fe^\top \rbM \ddtt \fe
        +
        \tfrac{1}{2}\fe^\top \ddtt{\rbM}\fe
        \\
        =&
        \fe^\top (
            -\rbVm \fe -\mm K \fe + \mv f
            -\rbd + \mysat(\rbu)
        )
        \\
        &
        +
        \tfrac{1}{2}
        \fe^\top \ddtt{\rbM}\fe
        \\
        =&
        -
        \fe^\top \mm K \fe 
        +
        \fe^\top (
            \mv f+\mysat(\rbu)
            -\rbd
        )
        \\
        &
        +
        \tfrac{1}{2}
        \fe^\top(\ddtt{\rbM}- 2\rbVm)\fe
        \\
        \le&
        -\lambda_{\min}(\mm K)\norm{\fe}^2
        +
        \overline\tau_d\norm{\fe}
        +
        \fe^\top (\mv f+\mysat(\rbu))
        % \\
        % \le&
        % -\lambda_{\min}(\mm K)\norm{\fe}^2
        % +
        % \overline\tau_d\norm{\fe}
        % +
        % \fe^\top (\mysat(\rbu)-\rbu^*)
        .
    \end{aligned}
    \label{eq:lya1:dot1}
\end{equation} 
Therefore, invoking Lemma \ref{lem:stable:set}, it can be concluded that the filtered error $\fe$ is exponentially stable, with $\lim_{t\to\infty}\norm{\fe}=\tfrac{\overline\tau_d}{\lambda_{\min}(\mm K)}$, provided that the lumped system uncertainty $\mv{f}$ can be perfectly cancelled by the control input $\rbu$, neglecting the effect of input saturation.
Hence, the ideal control input $\rbu^*$ is defined as $-\mv f$.
However, $\rbu^*$ is not available in practice, since system dynamics encapsulated in $\mv f$ are unknown.

To overcome this issue, a DNN is employed to approximate $\rbu^*$.
Let $\NN:=\NN({\q}_n;\wth): \R^{l_0+1}\times\R^{\Xi}\to\R^{n}$ represent the DNN, where ${\q}_n\in\R^{l_0+1}$ is the DNN input vector, and $\wth\in\R^{\Xi}$ is the vector of trainable weights.
The architecture of $\NN({\q}_n;\wth)$ will be presented in Section \ref{sec:sub:NN definition}.
According to the universal approximation theorem for DNNs \cite{Kidger:2020aa}, $\NN({\q}_n;\wth)$ can approximate a nonlinear function, denoted as $\mv g(\cdot)$, with an ideal weight vector $\idealwth$ over a compact subset $\Omega_{NN}\in\R^{l}$ to within $\epsilon$-accuracy, such that $\sup_{{\q}_n\in\Omega_{NN}}\norm{ \NN({\q}_n;\idealwth) - \mv g(\cdot)} = \epsilon < \infty$.
% Furthermore, the theorem states that the norm of $\idealwth$ is bounded, \ie $\norm{\idealwth}\le \overline\wth<\infty$.
In this study, $\idealwth$ is considered a locally optimal solution, rather than a global one.

Accordingly, the ideal control law $\rbu^*$ can be approximated by the DNN with the ideal weight vector $\idealNN:=\NN({\q}_n;\idealwth)$ as follows:
\begin{align}
    \rbu^*=& \idealNN+\mv\epsilon
    ,
    \label{eq:control:ideal}
\end{align}
where $\mv\epsilon\in\R^n$ is the approximation error vector, bounded by $\norm{\mv\epsilon}\le \bar{\epsilon}$.
The control input $\rbu^*$ is estimated online by
\begin{align}
    \rbu =& \estNN,
    \label{eq:control:est}
\end{align}
where $\hat\NN:=\NN({\q}_n;\estwth)$, and  $\estwth$ is the estimated weight vector for $\idealwth$.

Substituting \eqref{eq:control:ideal} and \eqref{eq:control:est} into \eqref{eq:lya1:dot1}, the time derivative of $V_1$ can be rewritten as
\begin{equation}
    \ddtt{V}_1
    \le 
    -\lambda_{\min}(\mm K)\norm{\fe}^2
    +
    \overline\tau_d\norm{\fe}
    +
    \fe^\top (\mysat(\hat\NN)-\idealNN-\mv\epsilon)
    .
    \label{eq:lya1:dot2}
\end{equation}
Therefore, one can conclude that the result of \eqref{eq:lya1:dot1} can be obtained by adapting $\estwth$ to $\idealwth$ (\ie $\hat\NN\to\idealNN$), neglecting the effect of input saturation.
Note that the control input saturation has not yet been considered in the derivation of the ideal control input; this will be addressed in the subsequent section.

@@

The DNN architecture $\NN({\q}_n;\wth) := \NN_k$ can be recursively represented as follows:
\begin{equation}
    \NN_i :=
    \begin{cases}
        \wV_i^\top \act_i(\NN_{i-1}), 
        &
        i\in\{1,\dots ,k\},
        \\
        \wV_0^\top {\q}_n,
        &
        i=0
        ,
    \end{cases}
    \label{eq:DNN:architecture}
\end{equation}
where $\wV_i\in\R^{(l_i+1)\times l_{i+1}}$ is the weight matrix of the $i\textsuperscript{th}$ layer, and $\act_i: \R^{l_i}\to\R^{l_i+1}$ represents the activation function of the $i\textsuperscript{th}$ layer. 
The activation function is defined as $\act_i(\mv x)=\left(\sigma(x_1),\sigma(x_2),\cdots, \sigma(x_{l_{i}}), 1\right)^\top,\ \forall\mv x\in\R^{l_i}$, where $\sigma: \R\to\R$ is a nonlinear function, and the augmentation of $1$ is used to account for bias terms in the weight matrices. 
Notice that the output size of $\NN(\cdot)$ is the same as that of the control input $\rbu$ (\ie $l_{k+1}=n$).

>>>>

\author{
  Myeongseok Ryu, Donghwa Hong, and Kyunghwan Choi
\thanks{
    Myeongseok Ryu and Kyunghwan Choi are with the Cho Chun Shik Graduate School of Mobility, Korea Advanced Institute of Science and Technology (KAIST), Daejeon 34051, Korea (e-mail: dding\_98@kaist.ac.kr and kh.choi@kaist.ac.kr).

@@

%% Abstract
\begin{abstract}
  This study presents a constrained optimization-based neuro-adaptive control (CONAC) for uncertain Euler-Lagrange systems subject to weight and convex input constraints. 
  A deep neural network (DNN) is employed to approximate the ideal stabilizing control law while simultaneously learning the unknown system dynamics and addressing both types of constraints within a constrained optimization framework.
  The adaptation law of DNN weights is formulated to solve the defined constrained optimization problem, ensuring satisfaction of first-order optimality conditions at steady state.
  The controller's stability is rigorously analyzed using Lyapunov theory, guaranteeing bounded tracking errors and bounded DNN weights. 
  The proposed controller is validated through a real-time implementation on a 2-DOF robotic manipulator, demonstrating its effectiveness in achieving accurate trajectory tracking while satisfying all imposed constraints.
\end{abstract}

@@

\begin{itemize}
    \item $\otimes$ denotes the Kronecker product \cite[Chap. 7 Def. 7.1.2]{Bernstein:2009aa}.
    \item A vector and a matrix are denoted by $\mv{x}=[x_i]_{i\in\{1,\cdots,n\}}\in\R^n$ and $
        \mm A
        := 
        [a_{ij}]
        _{
            i\in\{1,\cdots,n\},j\in\{1,\cdots ,m\}
        }\allowbreak\in\R^{n\times m}
        $, respectively.
    \item $\myrow_i(\mm A)$ denotes the $i\textsuperscript{th}$ row of the matrix $\mm A\in\R^{n\times m}$. 
    \item For $\mm{A}\in\R^{n\times m}$, $\myvec(\mm A):=\left(\myrow_1(\mm A^\top),\cdots,\myrow_m(\mm A^\top)  \right)^\top\in\R^{nm}$ denotes the vectorization of $\mm A$.
    % \item $\mycol_i(A)$ and $\myrow_j(A)$ denote the $i\textsuperscript{th}$ column and $j\textsuperscript{th}$ row of $A\in\R^{n\times m}$, respectively.
    % \item $\myvec(A):= [\mycol_1(A)^\top  ,\cdots,\mycol_m(A)^\top  ]^\top   $ for $A\in\R^{n\times m}$.
    \item $\lambda_{\min}(\mm A)$ denotes the minimum eigenvalue of the matrix $\mm A\in\R^{n\times n}$.
    \item $\mm I_n$ denotes the $n\times n$ identity matrix, and $\mm 0_{n\times m}$ denotes the $n\times m$ zero matrix.
\end{itemize}

@@

More recently, neuro-adaptive control approaches have been introduced to approximate unknown system dynamics or entire control laws using neural networks (NNs) \cite{Farrell:2006aa}. 
NNs are well-known for their universal approximation property, which allows them to approximate any smooth function over a compact set with minimal error. 
Various types of NNs have been utilized in neuro-adaptive control, including simpler architectures like single-hidden layer (SHL) NNs \cite{Ge:2010aa, Yesildirek:1995aa} and radial basis function (RBF) NNs \cite{Liu:2013ab,Ge:2002aa}, as well as more complex models like deep NNs (DNNs) \cite{Patil:2022aa} and their variations. 
Conventionally, SHL and RBF NNs are often employed to approximate uncertain system dynamics or controllers due to their simplicity \cite{Esfandiari:2014aa,Esfandiari:2015aa,Yesildirek:1995aa,Gao:2006aa}.
However, since DNNs offer greater expressive power, making them more effective for complex system approximations \cite{Rolnick:2018aa}, variations of DNNs, such as long short-term memory (LSTM) networks for time-varying dynamics \cite{Griffis:2023aa} and physics-informed NNs (PINNs) for leveraging physical system knowledge \cite{Hart:2024aa}, have further extended the capabilities of neuro-adaptive control systems.
\MSRY{CNN도 넣을까요? Dixon교수님네도 저희 preprint논문 인용하셨더라구요}

A critical aspect of neuro-adaptive control is the NN weight adaptation law, which governs how NN weights are updated. 
Most studies derived these laws using Lyapunov-based methods, ensuring the boundedness of the tracking error and weight estimation error, thus maintaining system stability under uncertainty.

@@

Although both the existing methods for weight boundedness and input constraints have shown effectiveness, they come with significant limitations. 
Projection operators and modification techniques do not guarantee the optimality of the adapted weights. 
Moreover, auxiliary systems typically handle only simple forms of input constraints, such as individual input bound constraints, limiting their ability to address more complex, nonlinear constraints.

@@

The main contributions of this study are listed as follows:
\begin{itemize}
    \item A constrained optimization-based neuro-adaptive control (CONAC) is proposed, where the control problem is formulated as a constrained optimization problem. In this formulation, weight and convex input constraints are incorporated as inequality constraints, while minimizing a given objective function.
    % \item Any convex input constraints can be applied to the proposed CONAC, including input bound constraints, and norm constraints. Moreover, the proposed CONAC can handle any combination of these constraints, which can be useful in practical applications where multiple constraints need to be satisfied simultaneously.
    % \item Since the proposed CONAC approximate the entire ideal control law with no prior knowledge of the system dynamics, the system identification process can be avoided, which is often difficult and time-consuming to implement in practice. 
    \item The adaptation laws for DNN weights and Lagrange multipliers are systematically derived to solve the defined problem, using constrained optimization theory. These laws guarantee convergence to the first-order optimality conditions, specifically the Karush–Kuhn–Tucker (KKT) conditions, as described in \cite[Chap. 12 Thm. 12.1]{Nocedal:2006aa}.
    \item The proposed CONAC's stability is rigorously analyzed using Lyapunov theory, ensuring that the tracking error and DNN weights remain bounded. This analysis guarantees that the proposed CONAC can achieve stable adaptation online.
    \item The proposed CONAC is implemented in real-time on a 2-DOF robotic manipulator, demonstrating its effectiveness in achieving accurate trajectory tracking while satisfying all imposed constraints. The experimental results validate the theoretical findings and show the practical applicability of the proposed CONAC.
\end{itemize}

@@

The remainder of this paper is organized as follows. 
Section \ref{sec:Problem Formulation} presents the target system and control objective.
Section \ref{sec:ctrl design} introduces the proposed CONAC and the architecture of DNN in the controller. 
In Section \ref{sec:adap_laws}, the adaptation law is developed.
Section \ref{sec:stability} analyzes the stability of the proposed CONAC.
A real-time implementation of the proposed CONAC is presented in Section \ref{sec:sim}, where the proposed CONAC is applied to a 2-DOF robotic manipulator.
Finally, Section \ref{sec:conclusion} concludes the paper and discusses potential future work.
The candidates of convex input constraints are presented in Appendix \ref{sec:appen:cstr}.

@@

We consider an uncertain Euler-Lagrange system modeled as
\begin{equation}
    \rbM\ddq + \rbVm + \rbF + \rbG + \rbd
    =
    \mysat(\rbu)
    \label{eq:sys1}
\end{equation}
where $\q\in \R^n$ denotes the generalized coordinate, $\rbd$ represents the disturbance and $\rbu\in\R^n$ denotes the generalized control input. 
The terms $\rbM:=\rbM(\q)\in\R^{n\times n}$, $\rbVm:=\rbVm(\q,\dq)\in\R^{n\times n}$, $\rbF:=\rbF(\dq)\in\R^{n}$, and $\rbG:=\rbG(\q)\in\R^{n}$ represent the unknown inertia matrix, Coriolis/centripetal matrix, friction vector, and gravity vector, respectively.
The function $\mysat(\cdot):\R^n\to\R^n$ represents the inherent physical limitations of the actuators such that $\norm{\mysat(\rbu)}\le \overline\tau$, where $\overline\tau$ denotes the maximum norm of control input.
In this study, the saturation function $\mysat(\cdot)$ is assumed to be convex, which is practically common in many engineering systems, since the actuator limits are naturally convex, such as min/max torque limits, and bounds look like a box or ball in the input space.
To account for these limitations, it is essential to incorporate physically motivated constraints into the controller design.
Appendix \ref{sec:appen:cstr} introduces candidate constraints that can be applied to ensure compliance with these physical limitations.

@@

\begin{prop}
    The disturbance $\rbd$ is bounded, \ie $\norm{\rbd}\le \overline\tau_d$, where $\overline\tau_d\in\R_{\ge0}$.
    \label{prop:dis_bound}
\end{prop}

The control objective is to develop a neuro-adaptive controller that enables $\q$ to track a continuously differentiable desired trajectory $\qd:=\qd(t): \R_{\ge0} \to \R^n$, compensating for the unknown system dynamics while addressing the imposed constraints.
The desired trajectory $\qd(t)$ is supposed to satisfy the following assumption:
\begin{assum}
    The desired trajectory $\qd(t)$ is assumed to be bounded, \ie $\norm{\qd(t)}\le \overline{q}_d\in\R_{>0}$ , and available for designing a bounded control input.
    \label{assum:feasible}
\end{assum}

@@

The architecture of the proposed CONAC is illustrated in Fig.~\ref{fig:ctrl:diagram}, and consists of a DNN that functions as a neuro-adaptive controller, and a weight optimizer for the DNN.
Before proceeding to the controller design, some mathematical preliminaries are revisited in Section \ref{sec:sub:math preliminaries}. 
Section \ref{sec:sub:NAC} introduces the neuro-adaptive controller, followed by the DNN model in Section \ref{sec:sub:NN definition}. 
The constrained optimization-based weight adaptation law is presented in Section \ref{sec:adap_laws}.

@@

\begin{proof}
    The time derivative of $V$ can be rewritten as
    \begin{equation}
        \ddtt V
        \le
        -a_1 \norm{\mv{x}}^2 + a_2 \norm{\mv{x}}
        =
        -a_1
        \left(
            \norm{\mv{x}}-\tfrac{a_2}{2a_1}
        \right)^2
        +
        \tfrac{a_2^2}{4a_1}
        .
        \label{eq:lem:lya:dot}
    \end{equation}
    According to \eqref{eq:lem:lya:dot}, one can conclude that $\ddtt V$ is negative definite if $\norm{\mv{x}}>\tfrac{a_2}{a_1}$ holds.
    In other words, $\mv{x}$ remains within the bounded set $\{\mv{x}\mid\norm{\mv{x}}\le\tfrac{a_2}{a_1}\}$, since $\ddtt V$ is negative definite when $\mv{x}$ is outside the set.
\end{proof}

@@

First, a filtered tracking error $\fe\in\R^n$ is defined as 
\begin{equation}
    \fe := \ddtt{\mv e}-\mm\Lambda \mv e
    ,
    \label{eq:err:filtered}
\end{equation}
where $\mv e := \q - \qd$ denotes the tracking error, $\ddtt{\mv e} := \dq - \dqd$ represents the time derivative of the tracking error, and $\mm\Lambda\in\R^{n\times n}_{>0}$ is a user-designed filtering matrix.
Since \eqref{eq:err:filtered} is stable system, it implies that $\mv e$ is bounded if $\fe$ is bounded.

Using $\fe$, the system dynamics \eqref{eq:sys1} can be rewritten as
\begin{equation}
    \rbM \ddtt\fe
    =
    -\rbVm \fe
    -\mm K \fe
    + \mv f
    -\rbd + \mysat(\rbu)
    ,
    \label{eq:sys:filtered}
\end{equation}
where $\mm{K}=\mm{K}^\top\in\R^{n\times n}_{>0}$ denotes an arbitrary unknown matrix and $
    \mv f:= \mv F
    \left(
        \q,\dq,\qd,\dqd,\ddqd
    \right)
    =
    \mm K \fe
    +\rbM
    \left(
        -\ddqd+\mm\Lambda\ddtt{\mv e}
    \right)
    +
    \rbVm
    \left(
        -\dqd+\mm\Lambda\mv e
    \right)
    -
    \rbF
    -
    \rbG
    \in\R^n
$ denotes the lumped system uncertainty.
% \color{red}
% The unknown matrix $\mm{K}$ is not available to design in practice, 
% \color{black}

Consider the Lyapunov function $V_1:= \tfrac{1}{2}\fe^\top \rbM\fe$. 
Invoking Property \ref{prop:skew}, the time derivative of $V_1$ is
\begin{equation}
    \begin{aligned}
        \ddtt{V}_1
        =&
        \fe^\top \rbM \ddtt \fe
        +
        \tfrac{1}{2}\fe^\top \ddtt{\rbM}\fe
        \\
        =&
        \fe^\top 
        \left(
            -\rbVm \fe -\mm K \fe + \mv f
            -\rbd + \mysat(\rbu)
        \right)
        \\
        &
        +
        \tfrac{1}{2}
        \fe^\top \ddtt{\rbM}\fe
        \\
        =&
        -
        \fe^\top \mm K \fe 
        +
        \fe^\top 
        \left(
            \mv f
            -\rbd
            +\mysat(\rbu)
        \right)
        \\
        &
        +
        \tfrac{1}{2}
        \fe^\top(\ddtt{\rbM}- 2\rbVm)\fe
        \\
        \le&
        -\lambda_{\min}(\mm K)\norm{\fe}^2
        +
        \overline\tau_d\norm{\fe}
        +
        \fe^\top (\mv f+\mysat(\rbu))
        % \\
        % \le&
        % -\lambda_{\min}(\mm K)\norm{\fe}^2
        % +
        % \overline\tau_d\norm{\fe}
        % +
        % \fe^\top (\mysat(\rbu)-\rbu^*)
        .
    \end{aligned}
    \label{eq:lya1:dot1}
\end{equation} 
Therefore, invoking Lemma \ref{lem:stable:set}, it can be concluded that the filtered error $\fe$ is exponentially stable, with $\lim_{t\to\infty}\norm{\fe}=\tfrac{\overline\tau_d}{\lambda_{\min}(\mm K)}$, provided that the lumped system uncertainty $\mv{f}$ can be perfectly cancelled by the control input $\rbu$, neglecting the effect of input saturation.
Hence, the ideal control input $\rbu^*$ is defined as $-\mv f$.
However, $\rbu^*$ is not available in practice, since system dynamics encapsulated in $\mv f$ are unknown.

To overcome this issue, a DNN is employed to approximate $\rbu^*$.
Let $\NN:=\NN({\q}_n;\wth): \R^{l_0+1}\times\R^{\Xi}\to\R^{n}$ represent the DNN, where ${\q}_n\in\R^{l_0+1}$ is the DNN input vector, and $\wth\in\R^{\Xi}$ is the vector of trainable weights.
The architecture of DNN $\NN({\q}_n;\wth)$ will be presented in Section \ref{sec:sub:NN definition}, later.
According to the universal approximation theorem for DNNs \cite{Kidger:2020aa}, $\NN({\q}_n;\wth)$ can approximate a nonlinear function, denoted as $\mv g(\cdot)$, with an ideal weight vector $\idealwth\in\R^\Xi$ over a compact subset $\Omega_{n}\in\R^{l_0+1}$ to within $\epsilon$-accuracy, such that $\sup_{{\q}_n\in\Omega_{n}}\norm{ \NN({\q}_n;\idealwth) - \mv g(\cdot)} = \epsilon < \infty$.
% Furthermore, the theorem states that the norm of $\idealwth$ is bounded, \ie $\norm{\idealwth}\le \overline\wth<\infty$.
In this study, the ideal weight vector $\idealwth$ is considered a locally optimal solution, rather than a global one.

Accordingly, the ideal control law $\rbu^*$ can be approximated by the DNN with the ideal weight vector $\idealNN:=\NN({\q}_n;\idealwth)$ as follows:
\begin{align}
    \rbu^*=& \idealNN+\mv\epsilon
    ,
    \label{eq:control:ideal}
\end{align}
where $\mv\epsilon\in\R^n$ is the approximation error vector, bounded by $\norm{\mv\epsilon}\le \bar{\epsilon}$.
The control input $\rbu^*$ is estimated online by
\begin{align}
    \rbu =& \estNN,
    \label{eq:control:est}
\end{align}
where $\hat\NN:=\NN({\q}_n;\estwth)$, and  $\estwth\in\R^\Xi$ is the estimated weight vector for $\idealwth$.

Substituting \eqref{eq:control:ideal} and \eqref{eq:control:est} into \eqref{eq:lya1:dot1}, the time derivative of $V_1$ can be rewritten as
\begin{equation}
    \ddtt{V}_1
    \le 
    -\lambda_{\min}(\mm K)\norm{\fe}^2
    +
    \overline\tau_d\norm{\fe}
    +
    \fe^\top 
    \left(
        -\idealNN
        -\mv\epsilon
        +\mysat(\hat\NN)
    \right)
    .
    \label{eq:lya1:dot2}
\end{equation}
Therefore, one can conclude that the result of \eqref{eq:lya1:dot1} can be obtained by adapting $\estwth$ to $\idealwth$, \ie $\hat\NN\to\idealNN$, neglecting the effect of input saturation.
Note that the control input saturation has not yet been considered in the derivation of the ideal control input; this will be addressed in the subsequent section.

@@

The DNN architecture $\NN({\q}_n;\wth) := \NN_k$ can be recursively represented as follows:
\begin{equation}
    \NN_i :=
    \begin{cases}
        \wV_i^\top \act_i(\NN_{i-1}), 
        &
        i\in\{1,\dots ,k\},
        \\
        \wV_0^\top {\q}_n,
        &
        i=0
        ,
    \end{cases}
    \label{eq:DNN:architecture}
\end{equation}
where $\wV_i\in\R^{(l_i+1)\times l_{i+1}}$ is the weight matrix of the $i\textsuperscript{th}$ layer, and $\act_i: \R^{l_i}\to\R^{l_i+1}$ represents the activation function of the $i\textsuperscript{th}$ layer. 
The activation function is defined as $\act_i(\mv x)=\left(\sigma(x_1),\sigma(x_2),\cdots, \sigma(x_{l_{i}}), 1\right)^\top,\ \forall\mv x\in\R^{l_i}$, where $\sigma: \R\to\R$ is a nonlinear function, and the augmentation of $1$ is used to account for bias terms in the weight matrices. 
Notice that the output size of $\NN(\cdot)$ is the same as that of the control input $\rbu$, \ie $l_{k+1}=n$.
